---
title: "Stats 101"
author: "Steph Locke"
date:  "`r Sys.Date()`"
output: revealjs::revealjs_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, dev="svg",
     out.height = "500px",  strip.white=TRUE
     ,collapse = TRUE, dev.args = list(bg="transparent")
)
library(ggplot2)
library(vegalite)
library(DT)
library(readxl)
library(data.table)

vegalite <- vegalite() %>%
    cell_size(700, 500) %>%
  legend_color(orient="left")
#vegalite <- vegalite()

ggopts<-theme_bw(base_size=24)+theme(
    panel.background= element_rect(fill="transparent", colour="transparent"),
    plot.background= element_rect(fill="transparent", colour="transparent"),
    panel.border = element_rect(fill = NA, 
                colour = NA), 
    axis.ticks = element_line(colour = "white"),
    text=element_text(colour="white")
)

Mode <- function(x) {
ux <- unique(x)
ux[which.max(tabulate(match(x, ux)))]
}

myFile<-"Predicting Age.xlsx"
```

# Aim

## Aim

 - Predict the ages of SQL Saturday Exeter attendees

# Our data

## Our data
We have the results from our survey.
```{r stats101-1}
basedata<-setDT(read_excel(myFile))
results<-basedata[Age<100&Age>Experience,]
```
Our sample has `r nrow(basedata)` records but with basic data cleansing we will be working with `r nrow(results)`.

## Results
```{r eval=FALSE }
DT::datatable(results)
```

```{r echo=FALSE }
DT::datatable(results,style="bootstrap",options=list(dom = "lftir" ))
```

## Age Distribution
```{r stats101-2}
ggplot(results)+ggopts+
  aes(x = Age, y=..count..)+
  geom_histogram(bins = 10, fill="white", colour="white")
```

## Experience Distribution
```{r stats101-3}
vegalite %>%
  add_data(results) %>%
  encode_x("Experience","quantitative") %>%
  encode_y("*", "quantitative", aggregate="count") %>%
  bin_x(maxbins=10) %>%  mark_bar()
```

## Field Distribution
```{r stats101-4}
vegalite %>%
  add_data(results) %>%
  encode_x("Field","ordinal") %>%
  encode_y("*", "quantitative", aggregate="count") %>%
  mark_bar()
```


## All data
```{r stats101-5}
vegalite %>%
  add_data(results) %>%
  encode_x("Experience","quantitative") %>%
  encode_y("Age", "quantitative") %>%
  encode_color("Field", "nominal") %>%
  mark_point()
```


# Sampling

```{r stats101-6}
testPer<-.75
inTest<-sample(1:nrow(results),nrow(results)*testPer)
training<-results[inTest,]
holdout<-results[!inTest,]
```

## Training
```{r stats101-7}
vegalite %>%
  add_data(training) %>%
  encode_x("Experience","quantitative") %>%
  encode_y("Age", "quantitative") %>%
  encode_color("Field", "nominal") %>%
  mark_point()
```

## Holdout
```{r stats101-8}
vegalite %>%
  add_data(holdout) %>%
  encode_x("Experience","quantitative") %>%
  encode_y("Age", "quantitative") %>%
  encode_color("Field", "nominal") %>%
  mark_point()
```

# One size fits all

## One size fits all
We could take some measure of central tendency to predict the age of attendees.

```{r stats101-9}
averages<-training[,.(Mean=floor(mean(Age))
                      ,Median=floor(median(Age))
                      ,Mode=Mode(Age)
                      )]
knitr::kable(averages)
```

## Results
```{r stats101-10}
holdout[,colnames(averages):=averages]
holdout.m<-melt(holdout, measure.vars = c("Age",colnames(averages)))
vegalite %>%
   add_data(holdout.m) %>%
  encode_x("Name", "ordinal") %>%
  encode_y("value", "quantitative") %>%
  encode_color("variable", "nominal")%>%
  mark_point()
```

## Assessing results
```{r stats101-11}
holdout.lse<-melt(holdout, measure.vars = colnames(averages))
holdout.lse[,Error:=(Age-value)^2]
knitr::kable(holdout.lse[,.(LSE=sum(Error)), variable][order(LSE)])
```

# Line of best fit
$y=mx+c$

```{r stats101-12}
expLM<-lm(Age~Experience, training)
summary(expLM)
```

## Model
```{r stats101-13}
training[,expLMres:=expLM$fitted]
ggplot(training, aes(x=Experience, y=Age))+
  geom_point()+
  geom_line(aes(y=expLMres),colour="blue")+
  theme_minimal()
```

## Results
```{r stats101-14}
holdout[,expLMres:=predict(expLM,holdout)]
holdout.m<-melt(holdout, measure.vars = c("Age","expLMres"))
vegalite %>%
   add_data(holdout.m) %>%
  encode_x("Name", "ordinal") %>%
  encode_y("value", "quantitative") %>%
  encode_color("variable", "nominal")%>%
  mark_point()
```

## Assessing results
```{r stats101-15}
holdout.lse<-melt(holdout, measure.vars = c("expLMres",colnames(averages)))
holdout.lse[,Error:=(Age-value)^2]
knitr::kable(holdout.lse[,.(LSE=sum(Error)), variable][order(LSE)])
```

# Complex linear regression
## Complex linear regression
- $y=m_1 x_1 +...+ m_n x_n +c$
- Contrasts

```{r stats101-16}
fieldLM<-lm(Age~Experience + Field, training)
summary(fieldLM)
```

## Model
```{r stats101-17}
training[,fieldLMres:=fieldLM$fitted]
ggplot(training, aes(x=Experience, y=Age, group=Field, colour=Field))+
  geom_point()+
  geom_line(aes(y=fieldLMres, group=Field),colour="blue")+
  facet_wrap(~Field)+
  theme_minimal()
```

## Results
```{r stats101-18}
holdout[,fieldLMres:=predict(fieldLM,holdout)]
holdout.m<-melt(holdout, measure.vars = c("Age","fieldLMres"))
vegalite %>%
   add_data(holdout.m) %>%
  encode_x("Name", "ordinal") %>%
  encode_y("value", "quantitative") %>%
  encode_color("variable", "nominal")%>%
  mark_point()
```

## Assessing results
```{r stats101-19}
holdout.lse<-melt(holdout, measure.vars = c("fieldLMres","expLMres",colnames(averages)))
holdout.lse[,Error:=(Age-value)^2]
knitr::kable(holdout.lse[,.(LSE=sum(Error)), variable][order(LSE)])
```

